---
title: "Decomposing Efficiency - (Data Envelopment Analaysis, VRS)"
author: '**Blessings Petro**'
date: "`r Sys.Date()`"
output: pdf_document
---

We will use the BCC DEA method. Assumptions are:

  - Variable Returns to scale
  - Decision-Making Units are operating at different Scales

First you install the a package to use to run the analysis. We will use the `Benchmark` package which we can install using the code below.
```{r}
# code below commented out on purpose

#install.packages("Benchmark", dependencies = TRUE)
```
--- 

After the package is installed, we can now move on to Analysis. In the code below, we called the necessary packages for the purposes of the analysis and for loading the data

```{r}
# calling the packages we'll use in the analysis

library(readxl) #for importing
library(writexl) #for exporting
library(lpSolveAPI) #for optimizing(Linear Programming)
library(ucminf) #misc, necessary for Benchmarking
library(quadprog) #misc, necessary for Benchmarking
library(knitr) #for tables
library(ggplot2) #for plots
library(Benchmarking) #for DEA


# loading the data

data= read_xlsx(
  "C:\\Users\\user\\Documents\\Data Clients\\Dyson Mponda\\Research Data-1.xlsx",
  sheet = "Sheet1",
  range = "A2:H82"
)
```
\ 

--- 

The data has been loaded and kept in a variable called `"data"`. To get an overview of the data:
```{r}
kable(head(data, 5))
```
\ 

\ 

## Performing the **DEA**

```{r fig.cap= "Boxplot Showing Distribution of Efficiency Scores", fig.height=3, fig.width=5}
# defining Inputs(x)
x_values= data.frame(data$`Seed (MK)`, data$`Labour (MK)`, data$`Pestcides and Herbicides (MK)`)

# defining Outputs(y)
y_values= data.frame(data$`Soya bean(quantity)`, data$Income)

# Perform DEA
dea_result <- dea(
  X = x_values,
  Y = y_values,
  RTS = "vrs", 
  ORIENTATION = "in"
)

# Extraction of the efficiency scores
efficiency_scores <- dea_result$eff

# Create a data frame for plotting
data$Efficiency <- efficiency_scores

# View results distribution
ggplot(data= data, aes(x= Efficiency))+
  geom_boxplot()
```

**NB:** output `Fertilizer` ommitted because of zeros, the model failed to optimize.


## Summary Statistics of the Efficiency Scores
```{r}
summary(data$Efficiency)
```


## Exporting the Data
```{r}
write_xlsx(data, "C:\\Users\\user\\Documents\\Data Clients\\Dyson Mponda\\DEA Scores.xlsx")
```


\newpage
---
**Objective 2: Tobit Regression Model**
---

$$
Y_i^* = \beta_0 + \beta_i X_i + \varepsilon_i
$$
$$
\varepsilon \sim N(0, \sigma^2_{\varepsilon})
$$
\[
Y_i = 
\begin{cases}
Y_i^* & \text{if } Y_i^* < b \\
b & \text{if } Y_i^* \geq b
\end{cases}
\]

where:\ 

  - $Y_i^*:$ the dependent/latent variable\ 
  - $\beta_0:$ the constant term\ 
  - $\beta_i:$ a vector of coefficients\ 
  - $X_i:$ a vector of independent variables(socio-economic and institutional factors)\ 
  - $\varepsilon_i:$ the error term\ 
  - $b:$ is the censoring threshold (=0.75)\ 

